\section{Ternary Representation: The Computational Encoding of Triple Equivalence}

\subsection{From Equivalence to Encoding}

The Triple Equivalence Theorem establishes that oscillation, category, and partition are mathematically identical. This identity demands a representation system that encodes three-dimensional structure natively. Binary representation, with its one-dimensional $2^k$ hierarchy, cannot do this without coordinate transformation. Ternary representation provides the natural encoding.

\begin{theorem}[Representational Necessity]
If oscillation = category = partition (triple equivalence), then the natural representation base is 3, not 2.
\end{theorem}

\begin{proof}
The triple equivalence yields three S-entropy coordinates $(S_k, S_t, S_e)$. Any position in this space requires specifying refinement along one of three axes. A ternary digit (trit) $t \in \{0, 1, 2\}$ maps directly:
\begin{align}
t = 0 &\leftrightarrow \text{refine } S_k \text{ (knowledge/oscillatory)} \\
t = 1 &\leftrightarrow \text{refine } S_t \text{ (time/categorical)} \\
t = 2 &\leftrightarrow \text{refine } S_e \text{ (entropy/partition)}
\end{align}
Binary representation would require encoding three values in two symbols, introducing artificiality. Ternary representation is structurally matched to the triple equivalence.
\end{proof}

\subsection{The Trit-Coordinate-Perspective Correspondence}

\begin{definition}[Trit-Coordinate-Perspective Mapping]
A single trit encodes three equivalent pieces of information:
\begin{center}
\begin{tabular}{cccc}
\textbf{Trit} & \textbf{S-Coordinate} & \textbf{Perspective} & \textbf{Physical Meaning} \\
\hline
0 & $S_k$ & Oscillatory & Frequency/phase \\
1 & $S_t$ & Categorical & State/distinction \\
2 & $S_e$ & Partition & Selection/aperture \\
\end{tabular}
\end{center}
\end{definition}

This mapping is not arbitrary but \textit{necessary}: it is the triple equivalence expressed in representational form.

\subsection{The $3^k$ Hierarchical Structure}

\begin{theorem}[$3^k$ Hierarchy]
At hierarchical depth $k$:
\begin{enumerate}
    \item There exist $3^k$ distinct cells in S-entropy space
    \item Each cell is addressed by a unique $k$-trit string
    \item The string encodes both position and the trajectory that reached it
\end{enumerate}
\end{theorem}

The hierarchy grows as:
\begin{center}
\begin{tabular}{cc}
\textbf{Depth $k$} & \textbf{Cells} \\
\hline
1 & 3 \\
2 & 9 \\
3 & 27 \\
6 & 729 \\
10 & 59,049 \\
\end{tabular}
\end{center}

Compare to binary: 6 bits encode only 64 values vs. 729 for 6 trits. Ternary is more information-dense because it matches the dimensionality of S-space.

\subsection{Trajectory-Position Duality}

\begin{theorem}[Address = Trajectory]
A ternary string simultaneously encodes:
\begin{enumerate}
    \item \textbf{Position}: Which cell in the $3^k$ hierarchy
    \item \textbf{Trajectory}: The sequence of refinement steps
    \item \textbf{Description}: Which perspective was used at each step
\end{enumerate}
These are not three separate pieces of information but three views of the same object.
\end{theorem}

This is the Decoupling Theorem in representational form: the path (trajectory) and the destination (position) are encoded identically. Finding and explaining are not distinguished at the representation level.

\subsection{Continuous Emergence}

\begin{theorem}[Continuous Emergence]
As $k \to \infty$, the discrete $3^k$ cell structure converges exactly to the continuous space $[0,1]^3$:
\begin{equation}
\lim_{k \to \infty} \text{Cell}(t_1, t_2, \ldots, t_k) = \mathbf{S} \in [0,1]^3
\end{equation}
An infinite ternary expansion specifies a unique point in the continuum.
\end{theorem}

This resolves the discrete-continuous duality:
\begin{itemize}
    \item \textbf{Categorical} (discrete): Finite-$k$ ternary strings
    \item \textbf{Oscillatory} (continuous): Infinite-$k$ limit
    \item \textbf{Partition} (both): The boundary between discrete cells
\end{itemize}

The three perspectives are limiting cases of ternary representation.

\subsection{Ternary Operations}

Boolean logic (AND, OR, NOT) is replaced by ternary primitives:

\begin{definition}[Ternary Primitives]
\begin{enumerate}
    \item \textbf{Projection} $\pi_i$: Extract refinements along axis $i$
    \begin{equation}
    \pi_i(t_1 t_2 \cdots t_k) = \{t_j : t_j = i\}
    \end{equation}
    
    \item \textbf{Completion} $\kappa$: Predict trajectory endpoint
    \begin{equation}
    \kappa(t_1 \cdots t_j) = t_1 \cdots t_j \cdot \hat{t}_{j+1} \cdots \hat{t}_k
    \end{equation}
    where $\hat{t}$ are predicted by categorical dynamics
    
    \item \textbf{Composition} $\circ$: Concatenate trajectories
    \begin{equation}
    (t_1 \cdots t_j) \circ (t'_1 \cdots t'_m) = t_1 \cdots t_j t'_1 \cdots t'_m
    \end{equation}
\end{enumerate}
\end{definition}

These operations ARE S-navigation:
\begin{itemize}
    \item Projection = reading one S-coordinate
    \item Completion = the Moon Landing Algorithm's trajectory prediction
    \item Composition = chaining navigation steps
\end{itemize}

\subsection{Hardware Instantiation}

Ternary logic instantiates in three-phase oscillators:

\begin{proposition}[Three-Phase Encoding]
Three oscillators with phases:
\begin{align}
\phi_0 &= 0 \\
\phi_1 &= 2\pi/3 \\
\phi_2 &= 4\pi/3
\end{align}
encode trits through phase leadership:
\begin{equation}
\text{trit} = i \Leftrightarrow \text{oscillator } i \text{ leads in phase}
\end{equation}
\end{proposition}

This is not exotic hardware---three-phase AC power systems are ubiquitous. The physical substrate for ternary computing already exists.

\subsection{Connection to the Gödelian Foundation}

The Gödelian residue $\mathcal{G}$ manifests in ternary representation:

\begin{proposition}[Finite Precision = Gödelian Residue]
For any finite $k$:
\begin{equation}
\text{Unaddressable portion} = \frac{1}{3^k} \times (\text{total S-space})
\end{equation}
As $k \to \infty$, this approaches zero, but never reaches it. The unaddressable portion IS $\mathcal{G} \equiv x$.
\end{proposition}

Finite ternary strings access $\infty - x$; the infinite extension required for exact addressing is structurally impossible for bounded formal systems.

\subsection{Connection to Categorical Memory}

Ternary representation IS categorical memory addressing:

\begin{enumerate}
    \item \textbf{Address format}: $k$-trit strings
    \item \textbf{Hierarchy}: $3^k$ cells at depth $k$
    \item \textbf{Navigation}: Trit-by-trit refinement
    \item \textbf{Clustering}: Similar trajectories $\to$ similar addresses $\to$ spatial proximity
\end{enumerate}

The memory controller navigates by extending ternary strings, with each precision-by-difference value determining the next trit.

\subsection{Connection to Poincaré Computing}

Ternary representation IS the phase space discretization:

\begin{enumerate}
    \item \textbf{Phase space}: $\mathcal{S} = [0,1]^3$
    \item \textbf{Discretization}: $3^k$ cells at depth $k$
    \item \textbf{Trajectories}: Sequences of trits
    \item \textbf{Recurrence}: Return to $\epsilon$-neighborhood = string similarity
\end{enumerate}

Poincaré recurrence in the $3^k$ hierarchy is string-matching: the trajectory returns when the trit sequence repeats (approximately).

\subsection{The Unified Picture}

Ternary representation unifies all components of the framework:

\begin{center}
\begin{tabular}{ll}
\textbf{Framework Component} & \textbf{Ternary Manifestation} \\
\hline
Triple Equivalence & Three trit values \\
S-entropy coordinates & Three refinement axes \\
$3 \times 3$ matrix & $3 \times 3$ trit pairs \\
Infinite recursion & Infinite trit extensions \\
Observation boundary $x$ & Unaddressable at finite $k$ \\
Categorical memory & $3^k$ address hierarchy \\
Poincaré computing & Trajectory = trit sequence \\
Moon Landing Algorithm & Trit completion prediction \\
\end{tabular}
\end{center}

\subsection{Summary: Ternary as Foundational}

Ternary representation is not a notational choice but a structural necessity:

\begin{enumerate}
    \item \textbf{Triple Equivalence $\Rightarrow$ Base 3}: Three equivalent perspectives mandate three-valued logic
    
    \item \textbf{S-Coordinates $\Rightarrow$ Trit-Axis Mapping}: Each trit refines one S-coordinate
    
    \item \textbf{Trajectory = Position}: Ternary strings encode both simultaneously
    
    \item \textbf{Continuous Emergence}: Infinite trits $\to$ continuous space
    
    \item \textbf{Finite Precision = Gödelian Residue}: The unaddressable IS $x$
    
    \item \textbf{Three-Phase Hardware}: Physical instantiation exists
\end{enumerate}

The entire S-entropy framework---epistemology, gas laws, memory, computing---reduces to ternary string manipulation. This is not reductionism but unification: the framework IS ternary dynamics in bounded $3^k$ phase space.

